Object-Oriented Internet - Azure Interoperability

### 2020-11-19 Latex/Manuscript.tex
Work out Markdown\40.OOITechnologyFeatures.md #28
- final latex text formatting
- fnal review of the .pdf

- fixes #28

\section{Cloud to Sensors Field Level Connectivity}\label{cloud-to-sensors-field-level-connectivity}

@@

\begin{figure}
      \centering
      \includegraphics[width=8cm]{../.Media/StrategyDomainModel.png}
      \caption{Strategy Domain Model}\label{figure.StrategyDomainModel}
\end{figure}

Unlike the above described solutions, the \textbf{embedded gateway} is not derived from the middleware concept. The domain model for this archetype is presented in the Fig.~\ref*{figure.StrategyDomainModel}. Promoting separation of concern design principle, the gateway functionality should be implemented as a self-contained software part embedded in the \textit{Networking} service of the \textit{Cyber-physical\ node}. Main functionality of this component is to transfer selected data between \textit{Cyber-physical\ network} using \textit{Networking} services of an existing \textit{Cyber-physical\ node} and \textit{Cloud-based\ front-end} using officially supported by the cloud vendor interconnection services.

@@

Based on the sessionless and session-oriented communication patterns examination against the IoT requirements \cite{mpostol2020} it could be concluded that the connectionless pattern better suites issues related to the assets mobility and traffic asymmetry that is characteristic for the application domains in concern. Additionally, to promote interoperability and address the demands of the M2M communication in the context of a multi-vendor environment the prototyping should use a framework that must be compliant with the OPC UA Part 14 PubSub spec \cite{RefWorks:doc:5d98837de4b055984c0eecf0}. According to proposed architecture (Fig.~\ref*{subs.architecture}) to implement the \textit{Embedded\ Gateway} as a composable part of the \textit{Cyber-physical\ Node} a library implementing \textit{Networking} functionality in compliance with mentioned above specification is a starting point for further development. Additionally it must be assumed that the library used to deploy \textit{Embedded\ Gateway} support dependency injection and be capable to compose an external part supporting Azure/pubsub gateway functionality. The composition process must be available without modification of the core code of an existing library. As a result the prototyping is to be limited to implementation of the \textit{Embedded\ Gateway} software part only.

@@

To promote interoperability and address the demands of the M2M communication in the context of a multi-vendor environment the prototyping should use a framework that must be compliant with the OPC UA Part 14 PubSub (Sect. \ref*{subs.architecture}) and support the \emph{Reactive Interoperability} (Sect. \ref*{introduction}) concept. A framework compliant with this requirements has been implemented as an open-source library named \textit{UAOOI.Networking.SemanticData} (SDK) under an umbrella of the project Object-Oriented Internet \cite{RefWorks:doc:5c66740ae4b081adf5804596}. The library is designed to be a foundation for developing application programs that are taking part in a message-centric communication pattern and interconnected using the reactive networking concept. The diagram in Fig.~\ref*{figure.UADataApplicationArchitecture} shows the relationship between the library \textit{SDK} and external parts composing any reactive networking application (\textit{Reactive\ Application}). The \textit{Reactive Application} is an aggregation of \textit{Producer} and \textit{Consumer} parts. They must support access to real-time process data, hence they are recognized as an extension of \textit{DataRepository} class. To implement the \textit{DataRepository} dedicated implementation of the \textit{IBinding} interface should be provided to create a bridge between CPS and an external row data represented by the \textit{LocalResources} class. A more in-depth description of the \textit{OOI Reactive Application} library enabling data exchange over a network using the reactive networking pattern is in \cite{mpostol2020}.

\begin{figure}
      \centering
      \includegraphics[width=8cm]{../.Media/UADataApplicationArchitecture.png}
      \caption{Reactive interoperability implementation}\label{figure.UADataApplicationArchitecture}
\end{figure}

@@

Many parts in the domain model presented in Fig.~\ref*{figure.UADataApplicationArchitecture} requires a polymorphic approach to implementation. To promote the polymorphic ready solution the following concepts have been adopted:

@@

\begin{itemize}
      \item \emph{IBindingFactory} - bidirectional data exchange with the underlying process,
      \item \emph{IConfigurationDataFactory} - the configuration data access,
      \item \emph{IMessageHandlerFactory} - pushing the \emph{Message} entities to/pulling from \emph{Distribution Channel},
      \item \emph{IEncodingFactory} - searching a dictionary containing value converters.
\end{itemize}

It is expected that the functionality implementation expressed by these interfaces is provided as independent external composable parts. The composition is accomplished at run time, and the effective application functionality depends essentially on reusable loosely coupled parts composed using the dependency injection software engineering and adaptive programming concept \todo{cite{RefWorks:doc:5d9796cbe4b0f66c52dccf04}}.

@@

The \textit{Consumer} and \textit{Producer} classes are derived from the \textit{DataRepository} (Fig.~\ref*{figure.UADataApplicationArchitecture}). The \textit{Consumer} uses the \textit{IBindingFactory} to gather the data recovered from the \emph{Message} instances pulled from a network. The received data may be processed or driven to any data destination, e.g. cloud-based front-end. The \textit{Producer} mirrors the \textit{Consumer} functionality and, after reading data from an associated source, populates the \emph{Message} using the gathered data.

@@

From the discussion covered by the Sect. \ref*{subs.architecture} we can infer that the interoperability of communicating parties requires establishing a semantic-context and security-context between them. If the received data is driven to any external data destination (e.g. cloud-based front-end) using out-of-band protocol semantic-context and security context must be defined for this interoperability relationship. There are a vast variety of possibilities for how to define these contexts. Anyway, the most common starting point is the configuration. The proposed approach to instantiate the application as a set of composable parts is very useful in this respect because it makes it possible to expand the core configuration functionality and adapt it to be helpful also in this case. Therefore, the semantic-context may be created based on the configuration using the implementation of the \textit{IConfigurationDataFactory} interface.

@@

This section describes an example implementation of an OPC UA PubSub to Azure gateway. This gateway is implemented as a composable part of the Reactive Networking Application (\textit{RxNetworking\ App}). The \textit{RxNetworking\ App} is an aggregation of \textit{Producer} and \textit{Consumer} entities derived from \textit{DataRepository}. They must provide interconnection to real-time process data, hence they are recognized as an extension of the \textit{DataRepository} class. \textit{AzureGateway} part fulfills the \textit{Consumer} role and uses out-of-band communication to push process data to the cloud.

\begin{figure}
      \centering
      \includegraphics[width=8cm]{../.Media/ImplementationDomainModel.png}
      \caption{Implementation Architecture}\label{figure.ImplementationDomainModel}
\end{figure}

@@

A generic domain model presenting relationship between the Azure \textit{Cloud-based\ front-end} and \textit{Cyber-physical\ node} attached to the \textit{Cyber-physical\ network} is presented in Fig. \ref*{figure.StrategyDomainModel}. As it was pointed out in the Sect. \ref*{subs.architecture} using OPC UA PubSub the aggregation of nodes by the network is loosely coupled, i.e. nodes can be added and removed from the network dynamically, and nodes may represent mobile data holding assets.

\begin{figure}
      \centering
      \includegraphics[width=\textwidth]{../.Media/ImplementationArchitecture.png}
      \caption{Implementation Architecture}\label{figure.ImplementationArchitecture}
\end{figure}

@@

The article provides an introductory understanding of the steps required to implement the \textit{Consumer} role of \textit{OOI\ Reactive\ Application}. The \textit{ReferenceApplication} is an example application of \textit{Semantic-Data} reactive networking based on OPC UA PubSub\cite{RefWorks:doc:5d98837de4b055984c0eecf0} \todo{cite OPC UA PubSub Main Technology Features}.

@@

Implementation of this interface is a basic step to implement \textit{Consumer} functionality. The \textit{DataRepository} represents data holding assets in the \textit{RxNetworking\ App} and, following the proposed architecture, the \textit{IBindingFactory} interface is implemented by this external part. It captures functionality responsible for accessing the process data represented by the \textit{LocalResources}. The \textit{LocalResources} represents the external part that has a very broad usage purpose. For example, it may be any kind of process data source/destination, and to name a few \textit{Raw\ Data}, \textit{OPC\ UA\ Address\ Space\ Management}, and \textit{Azure} services in this case.

@@

\subsubsection{\texorpdfstring{\textit{IConfigurationFactory} implementation}{IConfigurationFactory implementation}}\label{iconfigurationfactory-implementation}

Implementation of this interface is straightforward and based entirely on the library UAOOI.Configuration.Networking available as the NuGet package. In a typical scenario, this implementation should not be considered for further modification. The only open question is how to provide the name of the file containing the configuration of this role.

@@

The Azure supports the HTTP, AMQP, MQTT (Sec. \ref*{azure-main-technology-features}). On other hand the PubSub supports Ethernet, UDP, AMQP, MQTT {Sect. \ref*{subs.architecture}}. If the Ethernet or UDP has been selected to build interconnection based on PubSub direc interoperability between the Azure and PubSub is impossible because the mentioned protocols are not offered as the native communication services.

@@

\begin{figure}
      \centering
      \includegraphics[width=\textwidth]{../.Media/AzureGateway.png}
      \caption{Implementation Architecture}\label{figure.AzureGateway}
\end{figure}

>>>>

\section{Sensors to Cloud Field Level Connectivity}\label{cloud-to-sensors-field-level-connectivity}

@@

\begin{figure}
      \centering
      \includegraphics[width=8cm]{../.Media/StrategyDomainModel.png}
      \caption{Strategy Domain Model}\label{figure1.StrategyDomainModel}
\end{figure}

Unlike the above described solutions, the \textbf{embedded gateway} is not derived from the middleware concept. The domain model for this archetype is presented in the Fig.~\ref*{figure1.StrategyDomainModel}. Promoting separation of concern design principle, the gateway functionality should be implemented as a self-contained software part embedded in the \textit{Networking} service of the \textit{Cyber-physical\ node}. Main functionality of this component is to transfer selected data between \textit{Cyber-physical\ network} using \textit{Networking} services of an existing \textit{Cyber-physical\ node} and \textit{Cloud-based\ front-end} using officially supported by the cloud vendor interconnection services.

@@

Based on the sessionless and session-oriented communication patterns examination against the IoT requirements \cite{mpostol2020} it could be concluded that the connectionless pattern better suites issues related to the assets mobility and traffic asymmetry that is characteristic for the application domains in concern. Additionally, to promote interoperability and address the demands of the M2M communication in the context of a multi-vendor environment the prototyping should use a framework that must be compliant with the OPC UA Part 14 PubSub spec \cite{RefWorks:doc:5d98837de4b055984c0eecf0}. According to proposed architecture (Fig.~\ref*{figure1.StrategyDomainModel}) to implement the \textit{Embedded\ Gateway} as a composable part of the \textit{Cyber-physical\ Node} a library implementing \textit{Networking} functionality in compliance with mentioned above specification is a starting point for further development. Additionally it must be assumed that the library used to deploy \textit{Embedded\ Gateway} support dependency injection and be capable to compose an external part supporting Azure/pubsub gateway functionality. The composition process must be available without modification of the core code of an existing library. As a result the prototyping is to be limited to implementation of the \textit{Embedded\ Gateway} software part only.

@@

To promote interoperability and address the demands of the M2M communication in the context of a multi-vendor environment the prototyping should use a framework that must be compliant with the OPC UA Part 14 PubSub (Sect. \ref*{subs.architecture}) and support the \emph{Reactive Interoperability} (Sect. \ref*{introduction}) concept. A framework compliant with this requirements has been implemented as an open-source library named \textit{UAOOI.Networking.SemanticData} (SDK) under an umbrella of the project Object-Oriented Internet \cite{RefWorks:doc:5c66740ae4b081adf5804596}. The library is designed to be a foundation for developing application programs that are taking part in a message-centric communication pattern and interconnected using the reactive networking concept. The diagram in Fig.~\ref*{figure2.UADataApplicationArchitecture} shows the relationship between the library \textit{SDK} and external parts composing any reactive networking application (\textit{Reactive\ Application}). The \textit{Reactive Application} is an aggregation of \textit{Producer} and \textit{Consumer} parts. They must support access to real-time process data, hence they are recognized as an extension of \textit{DataRepository} class. To implement the \textit{DataRepository} dedicated implementation of the \textit{IBinding} interface should be provided to create a bridge between CPS and an external row data represented by the \textit{LocalResources} class. A more in-depth description of the \textit{OOI Reactive Application} library enabling data exchange over a network using the reactive networking pattern is in \cite{mpostol2020}.

\begin{figure}
      \centering
      \includegraphics[width=8cm]{../.Media/UADataApplicationArchitecture.png}
      \caption{Reactive interoperability implementation}\label{figure2.UADataApplicationArchitecture}
\end{figure}

@@

Many parts in the domain model presented in Fig.~\ref*{figure2.UADataApplicationArchitecture} requires a polymorphic approach to implementation. To promote the polymorphic ready solution the following concepts have been adopted:

@@

\begin{itemize}
      \item \emph{IBindingFactory} - bidirectional data exchange with the underlying process,
      \item \emph{IConfigurationDataFactory} - the configuration data access,
      \item \emph{IMessageHandlerFactory} - pushing the \emph{Message} entities to/pulling from the wire,
      \item \emph{IEncodingFactory} - searching a dictionary containing value converters.
\end{itemize}

It is expected that the functionality implementation expressed by these interfaces is provided as independent external composable parts. The composition is accomplished at run time, and the effective application functionality depends essentially on reusable loosely coupled parts composed applying the dependency injection software engineering and adaptive programming concept \todo{cite{RefWorks:doc:5d9796cbe4b0f66c52dccf04}}.

@@

The \textit{Consumer} and \textit{Producer} classes are derived from the \textit{DataRepository} (Fig.~\ref*{figure2.UADataApplicationArchitecture}). The \textit{Consumer} uses the \textit{IBindingFactory} to gather the data recovered from the \emph{Message} instances pulled from a network. The received data may be processed or driven to any data destination, e.g. cloud-based front-end. The \textit{Producer} mirrors the \textit{Consumer} functionality and, after reading data from an associated source, populates the \emph{Message} using the gathered data.

@@

From the discussion covered by the Sect. \ref*{subs.architecture} we can infer that the interoperability of communicating parties requires establishing a semantic-context and security-context between them. If the received data is driven to any external data destination (e.g. cloud-based front-end) using out-of-band protocol semantic-context and security-context must be defined for this interoperability relationship. There are a vast variety of possibilities for how to define these contexts. Anyway, the most common starting point is the configuration. The proposed approach to instantiate the application as a set of composable parts is very useful in this respect because it makes it possible to expand the core configuration functionality and adapt it to be helpful also in this case. Therefore, the semantic-context may be created based on the configuration using the implementation of the \textit{IConfigurationDataFactory} interface.

@@

\todo{replace RxNetworking by Reactive Application}
This section describes an example implementation of an OPC UA PubSub to Azure gateway. This gateway is implemented as a composable part of the Reactive Networking Application (\textit{RxNetworking\ App}). The \textit{RxNetworking\ App} is an aggregation of \textit{Producer} and \textit{Consumer} entities derived from \textit{DataRepository}. They must provide interconnection to real-time process data, hence they are recognized as an extension of the \textit{DataRepository} class. \textit{AzureGateway} part fulfills the \textit{Consumer} role and uses out-of-band communication to push process data to the cloud.

\begin{figure}
      \centering
      \includegraphics[width=8cm]{../.Media/ImplementationDomainModel.png}
      \caption{Implementation Architecture}\label{figure3.ImplementationDomainModel}
\end{figure}

@@

A generic domain model presenting relationship between the Azure \textit{Cloud-based\ front-end} and \textit{Cyber-physical\ node} attached to the \textit{Cyber-physical\ network} is presented in Fig. \ref*{figure1.StrategyDomainModel}. As it was pointed out in the Sect. \ref*{subs.architecture} using OPC UA PubSub the aggregation of nodes by the network is loosely coupled, i.e. nodes can be added and removed from the network dynamically, and nodes may represent mobile data holding assets.

\begin{figure}
      \centering
      \includegraphics[width=\textwidth]{../.Media/ImplementationArchitecture.png}
      \caption{Implementation Architecture}\label{figure4.ImplementationArchitecture}
\end{figure}

@@

\todo{review intro}The article provides an introductory understanding of the steps required to implement the \textit{Consumer} role of \textit{OOI\ Reactive\ Application}. The \textit{ReferenceApplication} is an example application of \textit{Semantic-Data} reactive networking based on OPC UA PubSub\cite{RefWorks:doc:5d98837de4b055984c0eecf0} \todo{cite OPC UA PubSub Main Technology Features}.

@@

Implementation of this interface is a basic step to implement \textit{Consumer} functionality. The \textit{DataRepository} represents data holding assets in the \textit{RxNetworking\ App} and, following the proposed architecture, the \textit{IBindingFactory} interface is implemented by this external part. It captures functionality responsible for accessing the process data represented by the \textit{LocalResources}. \todo{duplacted text} The \textit{LocalResources} represents the external part that has a very broad usage purpose. For example, it may be any kind of process data source/destination, and to name a few \textit{Raw\ Data}, \textit{OPC\ UA\ Address\ Space\ Management}, and \textit{Azure} services in this case.

@@

Implementation of the \textit{IConfigurationFactory} interface is straightforward and based entirely on the library UAOOI.Configuration.Networking available as the NuGet package. In a typical scenario, this implementation should not be considered for further modification. The only open question is how to provide the name of the file containing the configuration of this role.

@@

The Azure supports the HTTP, AMQP, MQTT (Sec. \ref*{azure-main-technology-features}). On other hand the PubSub supports Ethernet, UDP, AMQP, MQTT {Sect. \ref*{ooi-main-technology-features}}. If the Ethernet or UDP has been selected to build interconnection based on PubSub direc interoperability between the Azure and PubSub is impossible because the mentioned protocols are not offered as the native communication services.

@@

\begin{figure}
      \centering
      \includegraphics[width=\textwidth]{../.Media/AzureGateway.png}
      \caption{Implementation Architecture}\label{figure5.AzureGateway}
\end{figure}
